A Test of the Frequency Independence Assumption of Power System Coefficients used in Geomagnetically Induced Current Estimates

### 2019-07-16 paper.tex
Updates from Overleaf

One-second-cadence electric and magnetic field measurements from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.

@@

\section{Models and Methods}

@@

\section{Model Evaluation}

Out-of-sample estimation quality metrics for each model on day $D$ were determined by computing the average model coefficients using the remaining 34 days of data and then using these averaged coefficients to generate GIC time series and associated metrics on day $D$. This process was repeated to create a set of 35 out-of-sample prediction metrics for each model. In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the influence of transients that exist in the estimate at the start of the interval for the amount of time that the causal part of the impulse response (found by Fourier inversion of the transfer function) is non-zero and at the end for the amount of time that the acausal part of the impulse response is non-zero.

@@

The mean-squared-error for model $m$ on day $D$ is
\begin{equation*}
\mbox{MSE}_m = \frac{1}{N}\sum_{t=t_o}^{t_f} \big(GIC(t)-GIC_m(t)\big)^2 \,,
\end{equation*}
\noindent
where $GIC_m(t)$ is the model prediction of model $m$ given in Table~\ref{results} at time step $t$, $t_o=600$, and $t_f=85800$, so that sum is over one day of one-second-cadence measurements excluding the first and last 10~minutes.

@@

The signal-to-noise ratio on day $D$ at a given evaluation frequency is the ratio of of the smoothed power spectrum value of $GIC$ to the smoothed power spectrum value of the prediction error at that evaluation frequency. The smoothed spectrum was obtained by averaging the raw spectrum in the same frequency bands used for the regression. The average signal-to-noise ratio at a given evaluation frequency is the average of the 35 out-of-sample signal-to-noise ratios at that frequency.

@@

The PE and MSE ratios are shown in Table~\ref{results} and Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{results}, i.e., Model~1 has the lowest PE and Model~4 has the highest. Note that the large errors visible at the start and end of the time intervals are due to the transients discussed above and the first and last 10~minutes of each interval were omitted from metrics calculations.

\begin{table}
\caption{Out-of-sample model performance metrics for models $m=1-4$. MSE$_1$/MSE$_m$ is the ratio of the MSE for Model~1 to that of Model $m$. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
\centering
\begin{tabular}{l c c c c}
\hline
$m$\hspace{1em} Model & PE & 95\% CI & MSE$_1$/MSE$_m$ & 95\% CI\\
\hline
1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
2\hspace{1em} $G_E(t) = a(\omega)E_x(t) + b(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
3\hspace{1em} $G_{E'}(t) = a(\omega)E'_x(t) + b(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
\hline
\end{tabular}
\label{results}
\end{table}

@@

Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3. Note that electric and magnetic field data from Memambetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$. When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs. 0.78). (The \cite{Fujii2015} transfer function was obtained from the IRIS DMC EMTF database [\cite{Kelbert2011}]). Although not critical to the conclusions in this work, a preliminary investigation suggests several explanations. We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015}. After 2006, our calculated transfer functions differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%. In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used. Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer function parameters were computed using one-day segments of data and then averaged. We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged. This result is unexpected, and extensive successful verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code was MT software distributed with MT hardware), suggests the explanation is not due to an error in the code used her to compute transfer functions. We expect to consider this in a future investigation as it suggests that stack averaging can be a significant source of transfer function bias and note that the most likely explanation involves non-stationarity in the noise.

@@

To test the effect of different points on the line having different electric field spectra on the parameter estimates, we created a simulated GIC signal using electric field measurements made in Memambetsu (indicated with a superscript M) and Kakioka (superscript K), Japan:

@@

Although this proof-of-concept simulation shows how variations in the frequency content of the geoelectric field on the line in which GIC is measured can cause the data-estimated system coefficients to have a frequency dependence, it is very crude as it uses several approximations that were necessitated by the availability of data. The distance from Kakioka to Memambetsu is $\sim$1400~km, which is $\sim$14x that of the length of the power line ($\sim$100~km) on which GIC was measured, and the correlations between the electric field from the two sites are likely to be lower than the fields measured at points separated by a distance closer to the length of the power line. In addition, the discontinuous change in the electric field implicit in the model implies that there would be a part of the line that can have a non-zero net charge and thus is not physically realizable for typical conductors. A more realistic simulation would use electric field measurements that were both near a line on which GIC was measured and separated from each other by a distance that is comparable to the length of the line.

@@

>>>>

One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910 N, geodetic longitude: 144.189 E) were obtained from the Japan Meteorological Agency data portal.

@@

\section{Models and Methods}
\label{section:Models_and_Methods}

@@

\section{Model Evaluation}
\label{section:Model_Evaluation}

Out-of-sample estimation quality metrics for each model on day $D$ were determined by computing the average model coefficients using the remaining 34 days of data and then using these averaged coefficients to generate GIC time series and associated metrics on day $D$. This process was repeated to create a set of 35 out-of-sample prediction metrics for each model. In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the influence of transients that exist in the estimate at the start of the interval for the amount of time that the causal part of the transfer function impulse responses (computed by Fourier inversion of the interpolated transfer function components) is non-zero and at the end for the amount of time that the acausal part of the impulse responses are non-zero.

@@

The mean-squared-error for model $m$ on day $D$ is
\begin{equation*}
\mbox{MSE}_m = \frac{1}{N}\sum_{t=t_o}^{t_f} \big(GIC(t)-GIC_m(t)\big)^2 \,,
\end{equation*}
\noindent
where $GIC_m(t)$ is the model prediction of model $m$ at time step $t$, $t_o=600$, and $t_f=85800$, so that sum is over one day of one-second-cadence measurements excluding the first and last 10~minutes. The MSE ratio for model $m$ is MSE$_1$/MSE$_m$.

@@

The signal-to-noise ratio on day $D$ at a given evaluation frequency is the ratio of of the smoothed power spectrum value of $GIC(\omega_e)$ to the smoothed power spectrum value of the prediction error at that evaluation frequency. A smoothed spectrum was obtained by averaging the raw spectrum in the same frequency bands used for the regression described in Section~\ref{section:Models_and_Methods}. The average signal-to-noise ratio at a given evaluation frequency is the average of the 35 out-of-sample signal-to-noise ratios at that frequency.

@@

The PE and MSE ratios are shown in Table~\ref{results}. Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{results}, i.e., Model~1 has the lowest PE and Model~4 has the highest. Note that the large errors visible at the start and end of the time intervals are due to the transients discussed in \ref{section:Model_Evaluation} and the first and last 10~minutes of each interval were omitted from metrics calculations.

\begin{table}
\caption{Out-of-sample model performance metrics for Model $m$. MSE$_1$/MSE$_m$ is the MSE ratio of Model~1 with respect to that of Model $m$. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
\centering
\begin{tabular}{l c c c c}
\hline
$m$\hspace{1em} Model & PE & 95\% CI & MSE$_1$/MSE$_m$ & 95\% CI\\
\hline
1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
2\hspace{1em} $G_E(t) = a(\omega)E_x(t) + b(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
3\hspace{1em} $G_{E'}(t) = a(\omega)E'_x(t) + b(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
\hline
\end{tabular}
\label{results}
\end{table}

@@

Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3. Note that electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$. When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs. 0.78). (The \cite{Fujii2015} transfer function was obtained from the IRIS DMC EMTF database [\cite{Kelbert2011}]). Although not critical to the conclusions in this work, a preliminary investigation suggests several explanations. We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015}. After 2006, our calculated transfer functions differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%. In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used. Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer function parameters were computed using one-day segments of data and then averaged. We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged. This result is unexpected, and extensive successful verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code was MT software distributed with MT hardware), suggests the explanation is not due to an error in the code used her to compute transfer functions. We expect to consider this in a future investigation as it suggests that stack averaging can be a significant source of transfer function bias and note that the most likely explanation involves non-stationarity in the noise.

@@

To test the effect of different points on the line having different electric field spectra on the parameter estimates, we created a simulated GIC signal using electric field measurements made in Memanbetsu (indicated with a superscript M) and Kakioka (superscript K), Japan:

@@

Although this proof-of-concept simulation shows how variations in the frequency content of the geoelectric field on the line in which GIC is measured can cause the data-estimated system coefficients to have a frequency dependence, it is very crude as it uses several approximations that were necessitated by the availability of data. The distance from Kakioka to Memanbetsu is $\sim$1400~km, which is $\sim$14x that of the length of the power line ($\sim$100~km) on which GIC was measured, and the correlations between the electric field from the two sites are likely to be lower than the fields measured at points separated by a distance closer to the length of the power line. In addition, the discontinuous change in the electric field implicit in the model implies that there would be a part of the line that can have a non-zero net charge and thus is not physically realizable for typical conductors. A more realistic simulation would use electric field measurements that were both near a line on which GIC was measured and separated from each other by a distance that is comparable to the length of the line.

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/map.pdf}
\caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval. Note that the scales for the error time series differ from that for the top panel so that the error amplitudes are scaled up by a factor of 2.}
\label{predictions}
\end{figure}
